---
title: "collaborator analysis"
format: html
---

## Exercise: Explore, Clean, and Wrangle Data
```{r}
#| warning:: false
#| echo:: false

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```


read in the data
```{r}
lobster_traps <- read_csv("../data/Lobster_Trap_Counts_All_Years_20210519.csv")

colnames(lobster_traps)

summary(lobster_traps)
glimpse(lobster_traps)
head(lobster_traps)
```

## Convert missing values using `mutate()` and `na_if()`
```{r}

lobster_traps <- lobster_traps %>% 
  mutate(TRAPS = na_if(TRAPS, -99999))

unique(lobster_traps$TRAPS)
```

## `filter()` practice
```{r}
not_napl <- lobster_traps_clean %>% 
  filter(SITE != "NAPL")
head(not_napl)
```
* create and store a subset with lobsters at CARP or number of commercial trap floats is greater than 20
```{r}
lobster_traps_subset <- lobster_traps_clean %>% 
  filter(SITE == "CARP", TRAPS > 20)
head(lobster_traps_subset)
```

## Exercise: Data Visulization
```{r}
ggplot(data = lobster_traps,
       aes(x = TRAPS))+ 
  geom_histogram()+
  facet_wrap(~ YEAR)
```

* create a line graph of the number of total lobster commercial traps observed (y-axis) by year (x-axis) in the study, grouped by SITE.
```{r}

# first create a new data subset
lobsters_traps_summarize <- lobster_traps %>% 
  group_by(SITE, YEAR) %>% 
  summarize(TOTAL_TRAPS = sum(TRAPS, na.rm = TRUE))

# now create the line graph 
# line plot
ggplot(data = lobsters_traps_summarize, aes(x = YEAR, y = TOTAL_TRAPS)) +
    geom_line(aes(color = SITE))

# line and point plot
ggplot(data = lobsters_traps_summarize, aes(x = YEAR, y = TOTAL_TRAPS)) +
    geom_point(aes(color = SITE)) +
    geom_line(aes(color = SITE))

# bar graph that shows the amount of high and low fishing pressure of lobster commercial traps at each site from 2019-2021

# first create a new data subset

lobster_traps_fishing_pressure <- lobster_traps %>% 
  filter(YEAR %in% c("2019" , "2020" , "2021")) %>% 
  mutate(FISHING_PRESSURE = if_else(TRAPS > 8, true = "high", false = "low")) %>% 
           group_by(SITE, FISHING_PRESSURE) %>% 
           summarize(COUNT = n()) %>% 
           drop_na()
# create a bar graph using ggplot()
 
ggplot(data = lobster_traps_fishing_pressure,
       aes(x = SITE, 
           y = COUNT, 
           fill = FISHING_PRESSURE))+
  geom_col(position = "dodge")
 
```









